Staged Abstract Interpreters

### 2019-09-09 paper/20_prelim.tex
polish
typo

A monad is a type constructor @M[_]: * --> *@ with two operations, often called
@return@ and @bind@. Informally, @return@ wraps a value into a monad @M@, and
@bind@ unwraps a monadic value and transforms it into a new monadic value.
Pragmatically in Scala, we define the monad type class using trait @Monad@
(Figure \ref{fig:monad}), which declares the @pure@ \footnote{We elect to use
\texttt{pure} as the name, since \texttt{return} is a keyword in Scala and
\texttt{unit} is a built-in function in LMS.} and @flatMap@ operations. The
trait itself takes the monad type constructor @M[_]@ as an argument, which, as a
higher-kinded type, takes a type and returns a type. The method @pure@ promotes
values of type @A@ to values of type @M[A]@. The monadic @bind@ operation is
usually called @flatMap@ in Scala. It takes a monad-encapsulated value of type
@M[A]@, a function of @A => M[B]@, and returns values of type @M[B]@.

@@

Note that here the monadic object @List[_]@ encapsulates the data internally.
Therefore, it only exposes the simplified version of @flatMap@, where the monadic
value of @M[A]@ is not introduced as a function argument. The trait @MonadOps@ (Figure
\ref{fig:monadops}) defines the simplified versions of monadic operations that
are necessary for @for@-comprehension. The conversion between @Monad@ and
@MonadOps@ can be achieved by using the implicit design pattern.
In the rest of the paper, we use Scala's @for@-comprehension syntax and a few
monads and monad transformers such as @ReaderT@, @StateT@, and @SetT@ to
write our interpreters.
Monad transformers are type constructors of kind @(* --> *) --> (* --> *)@, which
take a monad as an argument and produce another monad. By using monad
transformers, we can combine multiple monads into a single one.  The
implementation of monads and monad transformers follows the one from Haskell.

>>>>

A monad is a type constructor @M[_]: * --> *@ with two operations, often called
@return@ and @bind@ \cite{Wadler:1992:EFP:143165.143169,
DBLP:journals/iandc/Moggi91}. Informally, @return@ wraps a value into a monad
@M@, and @bind@ unwraps a monadic value and transforms it into a new monadic value.
Pragmatically in Scala, we define the monad type class using trait @Monad@
(Figure \ref{fig:monad}), which declares the @pure@ \footnote{We elect to use
\texttt{pure} as the name, since \texttt{return} is a keyword in Scala and
\texttt{unit} is a built-in function in LMS.} and @flatMap@ operations. The
trait itself takes the monad type constructor @M[_]@ as an argument, which, as a
higher-kinded type, takes a type and returns a type. The method @pure@ promotes
values of type @A@ to values of type @M[A]@. The monadic @bind@ operation is
usually called @flatMap@ in Scala. It takes a monad-encapsulated value of type
@M[A]@, a function of @A => M[B]@, and returns values of type @M[B]@.

@@

Note that here the monadic object @List[_]@ encapsulates the data internally.
Therefore, it only exposes the simplified version of @flatMap@, where the monadic
value of @M[A]@ is not introduced as a function argument. The trait @MonadOps@ (Figure
\ref{fig:monadops}) defines the simplified versions of monadic operations that
are necessary for @for@-comprehension. The conversion between @Monad@ and
@MonadOps@ can be achieved by using the implicit design pattern.
In the rest of the paper, we use Scala's @for@-comprehension syntax and a few
monads and monad transformers such as @ReaderT@, @StateT@, and @SetT@ to
write our interpreters.
Monad transformers are type constructors of kind @(* --> *) --> (* --> *)@, which
take a monad as an argument and produce another monad. By using monad
transformers, we can combine multiple monads into a single one.  The
implementation of monads and monad transformers follows the ones from Haskell.
